Ensemble Stability of Biological Networks

### 2014-09-22 paper.tex
formatting

This description can be formalized as a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
 \begin{displaymath}
   \adj(G_X)_{ij} = \left\{
     \begin{array}{ll}
       1, & F_i \hbox{ depends on } x_j\\
       0, & F_i \hbox{ does not depend on } x_j
     \end{array}
   \right.
\end{displaymath}
A graph can be decomposed into a natural collection of \emph{modules} called strongly connected components. A strongly connected component of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}. The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc}.

@@

\section{Stability and robustness analysis of multi-component systems}
Suppose that $M$ is an $n \times n$ matrix with real coefficients and connectivity $\connectivity$. We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin, and having volume $r^\connectivity$.  We want to know the probability that, if we resample the elements $k_1, k_2, \ldots k_m$ of this matrix, the resulting matrix will correspond to a stable matrix which is given by a special case of \ref{eq:robustness}
\begin{align}\label{eq:condprobgen}
 R (stab, k_1, \ldots k_m) =
  \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{H^m} {dx'}_{k_1} \cdots {dx'}_{k_m}\,
    \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity)
    \mathbf{1}_{stab}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d)}
  {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)}
\end{align}

\todo{Generalize to resampling a random link and to resampling multiple edges.}

@@

The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of matrices whose connectivity is given by that digraph.  Say that the graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.  Label the vertices in such a way that $v_1, \ldots, v_{n_1}$ belong to $C_1$, $v_{n_1 + 1}, \ldots, v_{n_2}$ belong to $C_2$, etc.  Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.  This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.  Some simple examples of this are given by the graphs in \ref{fig:scc}.

>>>>

This description can be formalized as a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
 \begin{displaymath}
   \adj(G_X)_{ij} = \left\{
     \begin{array}{ll}
       1, & F_i \hbox{ depends on } x_j\\
       0, & F_i \hbox{ does not depend on } x_j
     \end{array}
   \right.
\end{displaymath}
A graph can be decomposed into a natural collection of \emph{modules} called strongly connected components. A strongly connected component of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}. The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc} along with their adjacency matrices and hierarchy diagrams (to be defined below).

The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of its adjacency matrix.  Say that the graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.  Label the vertices in such a way that $v_1, \ldots, v_{n_1}$ belong to $C_1$, $v_{n_1 + 1}, \ldots, v_{n_2}$ belong to $C_2$, etc.  Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.  This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.

@@

\section{Stability and robustness analysis of multi-component systems}
Suppose that $M$ is an $n \times n$ matrix with real coefficients and connectivity $\connectivity$. We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin, and having volume $r^\connectivity$.  We want to know the probability that, if we resample the elements $k_1, k_2, \ldots k_m$ of this matrix, the resulting matrix will correspond to a stable matrix.  This is a special case of \ref{eq:robustness}:
\begin{equation}\label{eq:condprobgen}
 R (\mathrm{stab}, k_1, \ldots k_m) =
  \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{H^m} {dx'}_{k_1} \cdots {dx'}_{k_m}\,
    \mathbf{1}_{stab} \mathbf{1}_{stab'}}
  {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)}
\end{equation}
where we made the abbreviations
\begin{align*}
\mathbf{1}_{stab} &= \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity) \\
\mathbf{1}_{stab'} &= \mathbf{1}_{stab'}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d)
\end{align*}
i.e. $\mathbf{1}_{stab'}$ corresponds to replacing $x_{k_1}, \ldots x_{k_m}$ with their primed counterparts.

While one can resample a fixed subset of the links, more typically we will be randomly selecting which links to resample with a uniform distribution on the links.  The result of this operation corresponds to averaging our quantity over subsets of links:
\begin{equation}
R (\hbox{avg stab}, m) =
\frac{1}{\binom{\connectivity}{m}}
\sum_{\{k_1, k_2, \ldots k_m\} \subset \{1, 2,, \ldots \connectivity\}}
R (\mathrm{stab}, k_1, \ldots k_m)
\end{equation}

@@
